Minigame characters:
Ladgerda
  Demographic: female; 46; Ulfen ethnicity (people of the north); Fighter, Viking subclass;
  (Ladgerda used to be an archer. Due to a serious accident long time ago, her left arm became weak and couldn't pull a bow.
  So she learnt how to use axe. She is still very brave, and fight like a viking.)
  Ladgerda is a very independet, competitive, and royal individual.

  Backstory (short):
  Ladgerda's husband Nicolas (Nico) was a captain of a mercenary detachment called New Normans.
  He was on a mission to escort a small group of viking sellters to find new land East of the Sanos Forest.
  However, Nico's group was attacked by gnomes and Shaoti clan members while they were traversing the forest.
  Nico's group was outnumbered. He surrendered and agreed to be Shaoti clan members' slave, so they would let other people go.
  Some survivors returned from Sanos Forest and shared this news with Ladgerda.
  They also told her the Shaoti clan members will probably sold Nico in Kaer Maga.

  As a result, Ladgerda decided to travel to Kaer Maga to find her husband.
  Because of the long and dangerous journey. Ladgerda asked her nephew __ (Ori's character to join; leave blank for now).


__ (Ori's character)
  Demographic: male; ; Human (mother is an Ulfen; father is a Chelaxian); ; ;
  __ is a very independent, knowledgable, and adventurous individual.

  Backstory (full version):
  __ is Aslog's youngest son. He was inspired to be a historian and an advanturer.
  He recently receivd a scholarship from the University of Korvosa to study history, which semester starts later this fall.
  He agreed to go with his anut Ladgerda to travel to Kaer Maga to find her husband.
  After that, he will venture south and go to Korvosa before the school year starts.

  __ 's father is a respected tailor in Roderick's Cove who own a clothing boutique. His father is also an avid reader.
  __ also enjoy reading books since he was little. __ could't bear the mundane task in his parents' clothing business.
  So he went to Riddleport looking for jobs. From bartender to assistant for a local crimelord,
  __ tried many jobs before he found his passion in learning/teaching history.
  He become a private history teacher for rich pirates/merchants who settled their homes in Riddleport, but too busy to accept formal education.
  At his early adulthood, __ met a sea merchant from Orison from tutoring (distant country southeast of Varisa).
  They dated for a few months, but the merchant decided to move on because she did not want to stay in Riddleport with __.
  Shortly afterward this episode, __ returned to Roderick's Cove and applied for college to further his education.
  The diverse population in Riddleport fascinated __,
  and made him want to learn more about the deep history and culture from around the Inner Sea Region.


Other important characters relevant to Ladgerda:
  Nicolas - Ladgerda's husband. A horselord. A captain of New Normans mercenary group.
    20 years ago, He met Ladgerda back in Velashu Upland when Nico and Ladgerda were both sergeants.
    10 years ago, Ladgerda was forced to retired due to a training accident, whereas Nico was promoted to captain.
    5 years ago, Nico fathered a child with Ladgerda named Henrik (after Ladgerda's father).
    2 years ago, Nico went on mission to discover new land in the south, hoping this would raise more money for his child.
    13 months ago, Nico found some good lands near Sanos Forest and convinced locals from the Upland to join him.
    Nico was believed to be captured by Shaoti clan members and sold into slavery.

  Aslog (61) - Ladgerda’s older half sister, __'s mother; Henrik’s first daughter.
    At age 11, Aslog went to Roderick's Cove with her bio mother after her parents' divorce.
    Aslog married Taylor at age 17. Gave birth to Helga 5 years later.
    She is kind and nurturing. She took care of Ladgerda when she was injured a few years ago.
    She Ladgerda, Ladgerda's son Henrik, and Ladgerda's bio mother Zoya to stay at her home when they have no place to go.

  Henrik Sr. - Ladgerda & Aslog's father. A Well respected soldier in the New Normans.
    He fought in the battle of Cryingleaf against giants where he was mortally wounded.
    He aspire Ladgerda to join the mercenary as well.

  Henrik Jr. - Ladgerda's son. 5 year old when Ladgerda left to find Nico.
    Aslog's older daughter Helga agreed to take care of Henrik.

  Taylor Gildersleeve (68) - Aslog's husband. Father of __ . Local tailor and owner of a clothing boutique.
    He is the distant relative of Jess Gildersleeve - port governor of Roderick's Cove.
    He loves reading. He taught Aslog and __ how to read.

  Helga Gildersleeve - __'s older sister, Taylor and Aslog's daughter.
    Helga is 39 (slightly younger than Ladgerda).
    Helga and her husband current run her parents' boutique.

Notes:
  New Normans is a fictional mercenary group I created.
    The group is located in southern Uplands to protect locals from raiders, but it also has chapters in Riddleport and Roderick's Cove.
    It could be replaced by any mercenary group formed predominately by Ulfens and nomadic people from the Velashu Uplands region.

  The character's age & date are estimates and subject to change.

Copyright (c) Yiwan Ye 2018 Copyright Holder All Rights Reserved.
